---
title: "Gene Expressions and Cancer"
author: "Kyler Kendall"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = TRUE, Warning = FALSE, fig.align = "center")
library(DataComputing)
library(ggplot2)
```

# Introduction

## Backgorund of Project

In the 1980s, the National Cancer Institute developed a set of 60 cancer cell lines, called NCI60. The original purpose of the project was for screening anti-cancer drugs. More than 41,000 probes were used on each of the 60 cell lines. The numbers in the NCI60 data set are how much a particular gene was expressed in one cell line. One way to tell how a gene is related to cancer is the differences in expression of that gene from one cancer type to another.

## Biology Talk

* What is a gene's expression level?

>* How is it calculated?
    * $log_2(T/R)$
>* What does the expression level tell us about the gene in a cancerous cell?
    * Upregulation
    * Downregulation
    * No change

# Analysis

## Let's get started!

```{r echo=FALSE, warning=FALSE}
NCI60 %>%
  head(10) %>%
  knitr::kable()
```

## 

```{r echo=FALSE, warning=FALSE}
NCI60cells %>%
   head(10) %>%
  knitr::kable()
```

## Clean Up of `NCI60`

```{r, warning=FALSE}
Narrow <- 
  NCI60 %>%
  tidyr:: gather(cellLine, expression, -Probe)
```

##

```{r}
Narrow %>%
  head(10) %>%
  knitr::kable()
```

## Onto `NCI60cells`

```{r}
CellTypes <-
  NCI60cells %>%
  select(cellLine, tissue) %>%
  mutate(cellLine = gsub("\\:", ".", as.character(cellLine)))
```

##

```{r}
CellTypes %>%
  head(10)%>%
  knitr::kable()
```

## Now Let's Join Them

```{r, message=FALSE, warning = FALSE}
Narrow <- 
  Narrow %>%
  inner_join(CellTypes)
```

## That's What We Want!

```{r,echo=FALSE, message=FALSE, warning = FALSE}
Narrow %>%
  head(10)%>%
  knitr::kable()
```

## Let's Pick a Gene

```{r, warning=FALSE}
Probe_TOP3A <-
  Narrow %>%
  filter(Probe=="TOP3A")
```

##

```{r}
Probe_TOP3A %>%
   head(10) %>%
  knitr::kable()
```

## Graphing the Data

```{r, fig.cap = "This graph shows the individual data points of the probe TOP3A in each tissue type.", echo=FALSE}
Probe_TOP3A %>%
  ggplot(aes(x = tissue, y = exp(expression))) +
  geom_jitter(width = 0.3) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## We can do better!

```{r, echo = FALSE}
SummaryStats_example <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn = mean(expression, na.rm = TRUE),
            se = sd(expression, na.rm = TRUE)/sqrt(n())) %>%
  mutate(top = mn + 1.96 * se,
         bottom = mn - 1.96 * se)
```

```{r, fig.cap = "This graph displays all the data of the probe TOP_3A put together. The red bar on the error bar displays where the mean of the data is.", warning=FALSE, echo=FALSE}
Probe_TOP3A %>%
  ggplot(aes(x = tissue, y = exp(expression))) +
  geom_jitter(width = 0.3) +
  geom_errorbar(data = SummaryStats_example,
                aes(x = tissue,
                    y = exp(mn),
                    ymax = exp(top),
                    ymin = exp(bottom)), width = 0.5) +
   geom_segment(data = SummaryStats_example,
                aes(x = as.numeric(tissue) - 0.1, 
                    y = exp(mn), 
                    xend = as.numeric(tissue) + 0.1, 
                    yend = exp(mn)), color = "red", size = 1) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## What Are Confidence Intervals?


# Discussion

## Here's What We Know
